\graphicspath{{Chapter/Figs/introduction/}}
\chapter{Introduction}

\section{Motivation}
Automation plays an essential role in the development of modern transport, as automation is the natural direction to take on in the seek of increased safety, efficiency and passenger comfort \cite{Lutin2018}. Autonomous racing provides a competition-driven framework for the exploration of autonomous driving, which incentivizes new innovations to take place. Thereby, racing often sets the starting point for new innovations that seek to revolutionize the entire industry \cite{Foxall91}. One example of such competition is \ac{fsd}.\footnote{\url{https://www.formulastudent.de/teams/fsd/}} \ac{fsd} challenges teams across the world to build cars that can autonomously drive around fixed tracks which are defined by differently colored cones. One car is racing at a time and is competing for the fastest lap time.\\
\pic{Layout of a \ac{fsd} track (Source: FSG21 Competition Handbook, p. 14, "Figure 2: Trackdrive")}{trackdrive}{1}
\\The problem of autonomous racing in this context can be split into three main parts: landmark detection and tracking, map generation and trajectory planning, and controlling the vehicle. The first step in autonomously driving a vehicle is to generate an abstract representation of its surroundings. In order to do this sensory input such as camera images, LIDAR data, and odometric input from an \ac{imu} is used to create and track landmarks in a virtual space and locate them relative to the vehicle. This task can be accomplished by \ac{slam} algorithms \cite{Singandhupe2019} and is not part of this thesis. Regarding the process of steering, the vehicle uses specific driving parameters such as desired velocity and steering angle to control the various actuators, e.g. motors, that move the vehicle. This problem is very similar to the controlling of non-autonomous manually driven vehicles, since the main difference is the driving parameters coming from sensors like the acceleration pedal and steering wheel in manual driving as opposed to the output of a processing pipeline in autonomous driving. This is also not part of this thesis. The problem that is left to solve is using the virtual space provided by the \ac{slam} to determine the driving parameters velocity and steering angle. This problem can be split into two parts: Map generation, which focuses on transforming information about landmarks into an abstract map of the racing track – and trajectory planning, which uses the abstract map to plan actions that will lead the vehicle to move along the track. This thesis looks at an extension of a classical algorithm for map generation which has been previously worked on and a novel machine learning approach to solve map generation and trajectory planning in one step and systematically compares these two approaches.\\



\section{Goals}
The framework for the implementation and application of the ideas presented in this thesis is set by Raceyard – a team from Kiel which is aminig to compete in \ac{fsd}. As of writing this thesis, a simplistic classical approach to map generation is used at Raceyard which imposes several problems that render the algorithm not yet usable in practice. For three of these problems this thesis suggests an improvement:

\begin{itemize}
    \item Robustness against the incorrect detection of the color of landmarks (misdetection), missing landmarks entirely (non-detection) and detection of landmarks twice or more with one detection being at the wrong place (over-detection). Using the current approach, only some misdetections can be automatically corrected. Any misdetection that cannot be corrected renders the resulting map completely unusable. Furthermore, non-detections are completely ignored, which leads to problems especially in narrow curves. Over-detections are handled like normal landmarks leading to wrong predictions as well.
    \item Using the certainty the \ac{slam} provides: The \ac{slam} assigns covariances representing the certainty in x- and y-direction to each landmark detected. This covariance is completely ignored by the current algorithm, although it could be beneficial to use.
    \item Runtime: The current approach takes orders of magnitudes too long to be used in realtime.
\end{itemize}
Another goal of this thesis is providing a profound comparison of the different characteristics of the machine learning- and classical algorithms used, especially with regard to their error resilience, precision and runtime.

\section{Related Work}
Many publications in the field of autonomous driving can be found; however, each of these works focus on key aspects that differ from this thesis in one or more ways. 

With regard to the classical approach to map generation, several techniques have been documented. The following papers apply a classical algorithm specifically to the problem of autonomous racing in \ac{fsd}. AMZ Driverless \cite{kabzan2019amz} as well as Andresen et al. \cite{andresen2020} focus on an architecture using an ordinary \ac{slam} in conjunction with a Delauney triangulation to do path planning. Zeilinger et al. \cite{zeilinger2017} as well as KIT19d \cite{nekkah2020} use an \ac{ekf}-\ac{slam} to derive the centerline for trajectory planning directly. Additionally, these papers do not adress machine learning as an alternative for path planning.\\
\\
In machine learning some approaches to autonomous racing can be found; however, none of those apply \ac{ml} to the problem of map generation and path planning in \ac{fsd} specifically. Dewing \cite{DewingNowTI}
used a \ac{cnn} to solve autonomous driving in a virtual racing game, while Dziubiński\footnote{\url{https://medium.com/asap-report/training-a-neural-network-for-driving-an-autonomous\%2drc-car-3906db91f3e}, accessed 19.02.2022} documented the use of a \ac{cnn} for steering a toy car in open terrain without cones to mark the path.\\
\\
One notable exception that applied machine learning to the problem presented in \ac{fsd} specifically is the work of Georgiev \cite{georgiev2019}. Georgiev implemented the \ac{mppi} by Williams et al. \cite{williams2016} in the Formula Student racing environment. \ac{mppi} uses a path integral over several possible trajectories to derive the best possible future trajectory in path planning. A neural network is used to train the parameters of the \ac{mppi}.\\
\\
To the knowledge of the author, no full \ac{ml} approach has been made specifically in the context of map generation in \ac{fsd}. Furthermore, no comparison to a classical approach in \ac{fsd} has yet been conducted. This work evaluates a modified classic heuristic algorithm in comparison to a \ac{ml} approach in the context of \ac{fsd} racing.

\section{Thesis Structure}
In the following chapter, foundations and technical background are explained surrounding the two approaches and autonomous racing in general.

Thereafter, in the third chapter, the details of the classical and \ac{ml} approach, as well as their implementation, are presented.

In the fourth chapter the approaches are evaluated and compared
and in the last chapter, the results are summarized and several improvement ideas and suggestions for future work are listed.

